Quantum Semiprime Factorization: Leveraging Grover's Algorithm for Efficient Prime Decomposition

### 2025-03-02 LaTeX/paper.tex
Updated grover's definitions and some methodology

% Keep the title short enough to fit on a single line if possible.
% Don't end it with a full stop (period).  Don't use ALL CAPS.
\title{Leveraging Grover's Algorithm for Efficient Prime Decomposition}

@@

  In quantum computing, SP (semiprime) factoring is typically performed 
  with Shor's algorithm, but it is possible to achieve the same results
  more efficientily via Grover's algorithm\cite{whitlock2023quantumfactoringalgorithmusing}.

  Grover's Algorithm is used to search for one (or many) given state(s)
  among some set of possible outputs. In this case, given some semiprime,
  Grover's algorithm may be used to find the prime factors.

@@

What are we doing, broadly?

@@

@@

Demonstration of quantum advantage, or at least determination of
the problem scale at which quantum advantage will pull ahead.

@@

Discuss existing methods of SP factoring- both quantum and classical.

Mention reference paper, QFT arithmetic, CUDA-Q docs, shor's, grover's.

\section{Methodology}

Quantum algorithms take advantage of superposition and entanglement,
allowing this form of computation to perform operations on many 
values simultaneously, rather than checking each value individually
as is necessary in classical algorithms.

  \subsection{Using Grover's Algorithm to Invert Functions}

@@

@@

>>>>

% Keep the title short enough to fit on a single line if possible.
% Don't end it with a full stop (period).  Don't use ALL CAPS.
\title{Quantum Semiprime Factorization: Leveraging Grover's Algorithm for Efficient Prime Decomposition}

@@

  In quantum computing, SP (semiprime) factoring is typically performed 
  with Shor's algorithm, but it is possible to achieve the same results
  more efficientily via Grover's algorithm\cite{quantum_factoring, grover}.

  Key terms: Semiprime, superposition, entanglement, RSA encryption, (add more).

@@

The practicality of applied quantum computing is a topic of much debate,
and there are not many examples of quantum computers being able to out-perform
their classical counterpart, yet.

Quantum algorithms have been in development at a theoretical level for
decades, but the application of these methods is still in early stages.
It is quite common to find code which implements a quantum algorithm like
Grover's or Shor's, but very often these programs are hard-coded, and only
work for a small handful of specific values\cite{shors_ibm}.

@@

  Quantum computers are a technology very much still under development. 
  As such, one of the major limitations for applied quantum computing is 
  the number of qubits required to perform a given task. If it can be verified 
  that SP factoring with Grover's algorithm uses significantly less qubits than
  Shor's, then the application of these techniques will be feasible far sooner.

@@

The goal of this research endeavor is to estimate the problem scaling point
at which some quantum algorithm may be able to out-perform our best classical
solution to the same problem.

Grover's algorithm has an incredibly wide range of potential use-cases,
and an intention for this paper is to provide a generalized reference point
for how Grover's algorithm may be used to invert any function.

@@

  \subsection{Grover's Algorithm}

  Grover's Algorithm is a quantum computing method 
  that can be used to search a database in $O(\sqrt{n})$ time 
  rather than the naive classical time complexity of $O(n)$\cite{grover}.

  Given a function $f(x)=y$, where $x$ is unknown (index, prime factors, 
  sum components, etc.), and $y$ is known (array value, semiprime/product, 
  sum, etc.), Grover's algorithm effectively takes the role of $f'(y)$, 
  allowing for a potential speedup in finding whatever input(s) to $f(x)$
  will return $y$.

  This speedup comes from the fact that Grover's algorithm requires 
  $O(\sqrt{y})$ iterations, each of which have a time complexity proportional
  to that of $f(x)$.

  \subsection{Shor's Algorithm}

  Description of Shor's algorithm.

  \subsection{Quantum Factoring Algorithm with Grover Search}

  S. Whitlock, et al. present a method of SP factoring with Grover's
  algorithm. The implementation in their paper is highly optimized,
  and modifies the target value in their circuit from some semiprime $N$,
  to $M$, a reduced number uniquely tied to $N$, which has prime factors
  $p$ and $q$.

  While this approach is admirable, it introduces a level of complexity 
  that may hinder a learner's understanding of the mechanics at play,
  so the implementation shown in section \ref{sec:Methodology} forgoes
  this abstraction from $N$ to $M$, and instead implements an oracle which
  searches directly for prime factors $a$ and $b$ of a semiprime $N$.

\section{Methodology} \label{sec:Methodology}

Quantum algorithms take advantage of superposition and entanglement,
which enables many methods of computation which are otherwise impossible
with a classical computer. For example, by placing a set of $n$ qubits 
(otherwise known as a qubit \textbf{register}) in superposition, that register
simultaneously represents every value which can be represented in $n$ bits,
until measured. Once measured, a register in uniform superposition will 
collapse into one of these states with equal likelihood.

In a given register $R_n$ with $n$ qubits, values are represented in binary,
such, for example, if some number $N=3$ were to be represented classically
with $n=4$ bits, it might be seen in binary as $0011$, and on a qubit
register that number may be represented similarly as $| 0011 \rangle$.

  \subsection{Grover's Algorithm for General Function Inversion}

  Mention unitary matrices.

@@

  Reference figure 1.

  Operations can be performed on a register in superposition, and these operations
  will affect each potential state individually. This property is utilized by
  Grover's Oracle to selectively target specific states which match the target
  value, and the diffuser is used to increase the probability of measuring the
  inputs which resulted in this marked target state.

  \begin{figure}[h]
  %\centering
  \includegraphics[width=7.9cm]{grover_inversion.png}
  \caption{General circuit diagram for Grover's Algorithm, used to find 
  the inputs $x, y$ for which $f(x,y)=z$, for some known $z$.}
  \label{FIGURA1}
  \end{figure}

  Include one figure showing the oracle, and one figure showing the diffuser.

@@

  Include a drawing of the circuit.
